### 2016-05-09 writing/latex/abstracteng.tex
add Anders' suggestions

A permutation is an arrangement of \(n\) objects.
Avoidance classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. In this thesis, the author provides sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, the author completely
Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
% of sets of the form \(231,m_1\) and \(231,m_2\)

>>>>

A permutation is an arrangement of \(n\) objects.
Permutation classes of classical permutation patterns
capture many interesting properties, such as
stack-sortability, and have links to many different combinatorial objects.
Mesh patterns are an extension of classical patterns that allow additional
restrictions to be placed on occurrences of the pattern.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. We provide sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
have the same number of avoiders at every length, we completely
Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
% of sets of the form \(231,m_1\) and \(231,m_2\)

### 2016-05-09 writing/latex/coincidence.tex
add Anders' suggestions

If the shading set \(D\) is non-empty we can modify definition of the directions
slightly
\begin{equation*}
    \begin{aligned}
        N &= \{(a,b+1),(a+1,b+1)\} \\
        E &= \{(a+1,b),(a+1,b+1)\} \\
        S &= \{(a,b),(a+1,b)\} \\
        W &= \{(a,b),(a,b+1)\}
    \end{aligned}
\end{equation*}
And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
\end{definition}

@@

\begin{definition}
Considering
only adding the ascent, as adding a descent is very similar.
Given a mesh pattern \(p = (\pi, R)\) \(\adda[D]{p}{(a,b)}\) gives a mesh
pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+2\) defined by
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
    b+j, b+j & \text{if } i = a+j, j\in\{1,2\}
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \{(a+1,b),(a,b),(a+1,b),(a+2,b),(a+1,b+2)\} \cup \bigcup_{(i,j)\in R}r((i,j)),
\end{equation*}
where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
        \{(i,j+2)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
        \{(i+2,j)\} & \text{if } i>a, j<b\\
        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}
\end{definition}

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by application of \PropositionRef{prop:dom1}. % repeated?

@@

>>>>

If the shading set \(D\) is non-empty we can modify the definition of the directions
slightly
\begin{equation*}
    \begin{aligned}
        N &= \{(a,b+1),(a+1,b+1)\} \\
        E &= \{(a+1,b),(a+1,b+1)\} \\
        S &= \{(a,b),(a+1,b)\} \\
        W &= \{(a,b),(a,b+1)\}
    \end{aligned}
\end{equation*}
And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
\end{definition}

@@

\begin{definition}
Considering
only adding the ascent, as adding a descent is very similar.
Given a mesh pattern \(p = (\pi, R)\), \(\adda{p}{(a,b)}\) gives a mesh
pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+2\) defined by
\begin{equation*}
    \pi^\prime(i) = \begin{cases}
    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
    b+j, b+j & \text{if } i = a+j, j\in\{1,2\}
\end{cases}
\end{equation*}
and
\begin{equation*}
        R^\prime = \{(a+1,b),(a,b),(a+1,b),(a+2,b),(a+1,b+2)\} \cup \bigcup_{(i,j)\in R}r((i,j)),
\end{equation*}
where \(r((i,j))\) is defined by
\begin{equation*}
    r((i,j)) =
    \begin{cases}
        \{(i,j)\} & \text{if } i<a, j<b\\
        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
        \{(i,j+2)\} & \text{if } i<a, j>b\\
        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
        \{(i+2,j)\} & \text{if } i>a, j<b\\
        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
    \end{cases}
\end{equation*}
\end{definition}

@@

All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
explained by application of \PropositionRef{prop:dom1}. By experimentation
we see that there are at least \(29\) coincidence classes, and all of these
coincidences are explained by this Proposition.

@@

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
    By \NoteRef{not:downcmesh},
    %  any of these occurrences is also an occurrence
    % of \(m_1\). This proves that every occurrence of \(m_2\) is also an
    % occurrence of \(m_1\) and therefore
     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.

### 2016-05-09 writing/latex/introduction.tex
add Anders' suggestions

\chapter{Introduction\label{cha:introduction}}
\section{What is a Permutation?}
In \emph{The Art of Computer Programming} \cite[p.~45]{Knuth:1997:ACP:260999} Knuth states,
\enquote{A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
objects in a row}. When considering permutations we can consider them as
occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). A
permutation \(\pi\) can be written in two-line notation
\begin{equation*}
\pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}
However, the most frequent notation used in computer science is
\emph{one-line notation}, in this form we drop the top line of the two line
notation and are left with the following.
\begin{equation*}
\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}

@@

For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements
(\(453\)) are order isomorphic to \(\perm{231}\).
This can be seen graphically below, the subsequence order isomorphic to \(\sigma\)
are highlighted.
\begin{equation*}
    \pattern{2,4,5}{2,4,1,5,3}{}
\end{equation*}

@@

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
of \(\pi\) in \(\tau\) with the additional restrictions on the relative
position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
no elements of \(\tau\) are allowed in the regions of the plot corresponding
to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

@@

Now we consider the plot for the permutation \(\sigma\). This permutation
avoids \(p\) since for every occurrence of the classical pattern \(\perm{2,1,3}\)
there is at least one point in one of the shaded boxes. For example, consider the
subsequence\(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
not the mesh pattern since the points with values \(4\) and \(2\) are in the
regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
\(p\). This is shown in the figure below.

>>>>

\chapter{Introduction\label{cha:introduction}}
\section{What is a Permutation?}
In \emph{The Art of Computer Programming} \cite[p.~45]{Knuth:1997:ACP:260999} Knuth states,
\enquote{A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
objects in a row}.
When considering permutations we can consider them as
occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). A
permutation \(\pi\) can be written in two-line notation
\begin{equation*}
\pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}
However,
This
% most frequent
notation
% used in computer science
is called
\emph{one-line notation}.
% in this form we drop the top line of the two line
% notation and are left with the following.
In this form we write the entries of the permutation in order, and get
\begin{equation*}
\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}

@@

For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements
(\(453\)) are order isomorphic to \(\perm{231}\).
This can be seen graphically below, the subsequence order isomorphic to \(\sigma\)
is highlighted.
\begin{equation*}
    \pattern{2,4,5}{2,4,1,5,3}{}
\end{equation*}

@@

We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
of \(\pi\) in \(\tau\) with the additional restrictions on the relative
position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
no elements of \(\tau\) are allowed in the regions of the plot corresponding
to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

@@

Now we consider the plot for the permutation \(\sigma\). This permutation
avoids \(p\) since for every occurrence of the classical pattern \(\perm{2,1,3}\)
there is at least one point in one of the shaded boxes. For example, consider the
subsequence \(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
not the mesh pattern since the points with values \(4\) and \(2\) are in the
regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
\(p\). This is shown in the figure below.
